---
title: "hw03-xubin-kuang"
author: "Xu-Bin Kuang"
date: "15/10/2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Packages}
# packages
library(readr)    # importing data
library(dplyr)    # data wrangling
library(ggplot2)  # graphics

# import data
teams <- read_csv("../data/nba2017-teams.csv")
```

```{r NBA Teams Ranked by Total Salary}
ggplot(teams, aes(x = reorder(team, salary), y=salary)) +
  geom_bar(stat='identity') +
  coord_flip() +
  labs(x = "Team", y = "Salary (in millions)", 
       title = "NBA Teams Ranked by Total Salary") +
  geom_hline(yintercept = mean(teams$salary), color='red')
```

The Warriors score the most total points with roughly 8,500 points in total. The average number of points for all other teams is around 7,700 points.
```{r}
ggplot(teams, aes(x = reorder(team, points), y=points)) +
  geom_bar(stat='identity') +
  coord_flip() +
  labs(x = "Team", y = "Total Points", 
       title = "NBA Teams Ranked by Total Points") +
  geom_hline(yintercept = mean(teams$points), color='red')
```

Finally, we *rank by team efficiency.* 
The Warriors and the Cavaliers lead the league in terms of effcienc.

```{r}
ggplot(teams, aes(x = reorder(team, efficiency), y=efficiency)) +
  geom_bar(stat='identity') +
  coord_flip() +
  labs(x = "Team", y = "Total Efficiency", 
       title = "NBA Teams Ranked by Total Efficiency") +
  geom_hline(yintercept = mean(teams$efficiency), color='red')
```

###Principal Components Analsysis (PCA). 
```{r}
# data frame with eigenvalues
pca_vars <- select(teams, points3, points2, free_throws, off_rebounds, def_rebounds, assists, steals,
                   blocks, turnovers, fouls)
pca <- prcomp(pca_vars, scale. = TRUE)
eigen <- data.frame(eigenvalue = round(pca$sdev^2,4), prop = round(pca$sdev^2 / sum(pca$sdev^2), 4))
eigen$cumprop <- cumsum(eigen$prop)
eigen
```
*Plot PC1 and PC2.*
```{r}
# create another df
pca_df = as.data.frame(pca$x)
rownames(pca_df) <- teams$team
# head(pca_df)

# plot of observations
ggplot(data = pca_df, aes(x = PC1, y = PC2, label = rownames(pca_df))) +
  geom_hline(yintercept = 0, colour = "gray65") +
  geom_vline(xintercept = 0, colour = "gray65") +
  geom_text(colour = "black", alpha = 0.8, size = 4) +
  ggtitle("PCA Plot (PCA1 and PCA2)")
#plot(pca$PC1, pca$PC2)
```

*Transform PC1.*
```{r}
transformed_PC1 <- 100 * ((pca_df$PC1 - min(pca_df$PC1)) / (max(pca_df$PC1) - min(pca_df$PC1)))
transformed_PC1
```

*Plot with new PC1 (scaled from 0 to 100).* 
GSW is clearly the top team based on PCA with the Sacramento Kings being the worst

```{r}
ggplot(teams, aes(x = reorder(team, transformed_PC1), y=transformed_PC1)) +
  geom_bar(stat='identity') +
  coord_flip() +
  labs(x = "Team", y = "First PC (scaled from 0 to 100)", 
       title = "NBA Teams Ranked by Scaled PC1")
```
