---
title: "post2-XuBin-Kuang"
author: "Xu-Bin Kuang"
date: "28/11/2017"
output: html_document
---
#Post 2: Mapping Violent Crimes in America

##Introduction
Map diagrams have become a popular medium to illustrate data in modern media. Compelling and vivid, choropleth maps are a great method of data presentation when conducting analysis of datasets with geometric significance. Through this post, I would like to utilize the 'maps' and 'ggplot2' packages to generate these maps. This post shows how to use ggplot to map a choropleth map, changing legend attributes, scale, color, titles and exporting the map into an image file.

Beyond making maps, the chosen topic is also one of importance. Mapping the violent crimes in America could potentially lead to observations that might affect policy. This post contains some elementary statistical analysis on the topic and a few hypotheses behind them.

###Motivation
"And I'm from the murder capital where they murder for capital" - Kanye West (From 'Murder to Excellence' by Jay-Z & Kanye West)

That infamous line by Kanye West alludes to his upbringing in Chicago and the rampant crime rates associated with the city. Upon hearing that line recently, I thought about the rates of violent crime in the country and how it would be an interesting topic of research. Through this post, I will not only be able to verify Kanye's claim about Chicago being the murder capital but I will also be able to map other violent crimes like rape and assault. My goals are to hopefully obtain interesting oberservations about the violent crime rates in the country and also familliarize myself with the "maps" package. 

##Content
```{r}
library(readr)
library(ggplot2)
library(maps)
library(dplyr)
```

###Data
This data set contains statistics, in arrests per 100,000 residents for assault, murder, and rape in each of the 50 US states in 1973. Also given is the percent of the population living in urban areas.
```{r}
URL <- "https://raw.githubusercontent.com/vincentarelbundock/Rdatasets/master/csv/datasets/USArrests.csv"
download.file(URL, destfile='crime_data.csv', method = 'curl')
crimeData <- read.csv('crime_data.csv')
colnames(crimeData)[1] <-"region"
head(crimeData)
```

###Correlation Between Total Violent Crimes and Urban Population
It seems intuitive to me that crime rates would reduce in urban environments. My justificaiton for that claim can be broken down into two arguments. Firstly, any given street is less likely to be empty in a higher density city. Considering how crime thrives on anonymity, higher density should result in fewer cases of crimes. 

Secondly, if an area has more citizens, it also means that it can also support a greater police force. Policing efficacy is heavily reliant on not only the ratio of cops to citizens but also the ratio of cops to land area. Therefore, it is safe to assume that a denser city would be a better policed city.

```{r}
#Total Number if Violent Crimes in Each State
crimeData$totalCrime = crimeData$Murder + crimeData$Assault + crimeData$Rape
crimeData$region = tolower(crimeData$region)
#Correlation between Total Crime and Percentage of Urban Population
corr <- round(cor(crimeData$totalCrime, crimeData$UrbanPop),2)
```

```{r ScatterPlot}
ggplot(crimeData, aes(x = crimeData$UrbanPop, y = crimeData$totalCrime)) + 
  geom_point() + 
  geom_smooth() + 
  labs(x = "Percentage of Urban Population(%)", y = "Number of Violent Crimes", title = "Number of Violent Crimes vs Urban Population") +
  geom_text(x = 85, y = 10, label = paste("r =", corr)) + 
  theme(plot.title = element_text(hjust = 0.5))
```

From the scatterplot and correlation coefficient, it is evident that there isn't a linear correlation between Urban Population and Number of Violent Crimes. This is expected as the data set has many outliers. This is because there are many different factors like economic development and state legislaiton that affects crime rates across different states in the US that our scatter plot is unable to account for.

One observation that stood out to me was the steep initial increase in crime as density rises from rural to urban. My initial thoughts on this was that, because crime relies on anonymity, crimes do not happen as often in rural areas because the likelihood of being identified is high. 

More research and data is required to make any claims and observations regarding the phenomenon of crime in the US and it's relation to state density.

##Murder Capital of America?

```{r}
topMurder <- crimeData$region[crimeData$Murder == max(crimeData$Murder)]
topRape <- crimeData$region[crimeData$Rape == max(crimeData$Rape)]
topAssault <- crimeData$region[crimeData$Assault == max(crimeData$Assault)]
```


Let's begin the mapping process by plotting a basic map of the states in America. This can be accessed by the built-in "state" dataset in the maps package.
```{r}
usStates <- map_data("state")
head(usStates)
```

Maps can be plotted using the using the polygon geom (using geom_ploygon()). geom_ploygon() essentially draws lines between points and "closes them up"(i.e. draws a line from the last point back to the first point).

It's default is a black fill with no line color. I've made the lines white for clarity.

```{r}
#Generates a basic map of 
baseMap <- ggplot(data = usStates) + 
  geom_polygon(aes(x = long, y = lat, group = group), color = "white")
baseMap
```

In order to start plotting information into our map, the datasets need to be merged by a common variable(I used 'region'). Next, change the fill of the map to match the density of the variable you would like plotted. There are many ways to make chloropleth maps in R. Through my research and experience, I found this to be the easiest.


```{r}
#Merging DataSets
crimeData$region = tolower(crimeData$region)
mapData <- inner_join(usStates, crimeData, by = 'region')

#Map of Murder Crimes in America
mmap <- baseMap + geom_polygon(data = mapData, aes(x = long, y = lat ,group = group,fill = Murder),color = "white") + labs(title= "Murder in America") + theme(plot.title = element_text(hjust = 0.5)) + scale_fill_gradient(low = "white", high = "red")

mmap
```

Now that we have the desired map, we can do the same for Rape and Assault in America.

```{r}
#Map of Rape Crimes in America
rmap <- baseMap + geom_polygon(data = mapData, aes(x = long, y = lat ,group = group,fill = Rape),color = "white") + labs(title= "Rape in America") + theme(plot.title = element_text(hjust = 0.5)) + scale_fill_gradient(low = "white", high = "blue")

rmap
```

```{r}
#Map of Assault Crimes in America
amap <- baseMap + geom_polygon(data = mapData, aes(x = long, y = lat ,group = group,fill = Assault),color = "white") + labs(title= "Assault in America") + theme(plot.title = element_text(hjust = 0.5)) + scale_fill_gradient(low = "white", high = "orange")

amap
```

##Exporting the Map
Finally, we will now look into exporting the maps that we have created. A quick and simple way to do this is via the PDF function.

```{r eval = FALSE}
pdf("mymap.pdf")
print(p2)
dev.off()
```

##Conclusion

The maps generated turned out better than I had expected. Solely using intuition and observing the maps, some interesting questions arise. Why are stats with high murder raters predominantly concentrated in the East Coast? What factors might lead to this? Is it surprising that Nevada would have the highest rates of reported rape crimes? What makes assault crimes so rampant in North Carolina? 

Simply by looking at the maps, one can come up with many interesting topics for research. While I am unable to provide a definite answer to justify these findings, I am still confident that these choropleth maps serve as a great data visualization tool.

I hope that I have broken down the process of generating maps in R in a simple and readable manner. I highly encourage generating maps of your own as it introduces a new perspective. 

##References
  * Density and Crime - https://www.theatlantic.com/national/archive/2011/05/density-and-crime/238944/
  * How to make choropleth maps with R - http://bl.ocks.org/prabhasp/raw/5030005/
  * Mapping with ggplot: Create a nice choropleth map in R - http://rforpublichealth.blogspot.com/2015/10/mapping-with-ggplot-create-nice.html
  * Package 'choroplethr' CRAN.R-project.org - https://cran.r-project.org/web/packages/choroplethr/choroplethr.pdf
  * Introduction to visualizing spatial data in R - https://cran.r-project.org/doc/contrib/intro-spatial-rl.pdf
  * World Map with r? - https://stackoverflow.com/questions/20079308/world-map-with-r
  * Crime Trends:1990-2016 -https://www.brennancenter.org/publication/crime-trends1990-2016
  


